---
title: "Mini-Project 2"
author: "Dejia Tang, Zicong Wei"
date: "October 31, 2017"
output: html_document
---

```{r message=FALSE, echo=FALSE, warning=FALSE}
library(tidyverse)
library(ggthemes)
library(ggrepel)
```

```{r, echo=FALSE}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```

## Data Wrangling
```{r message=FALSE, echo=TRUE, warning=FALSE}

# Create df_candidate by joining candidates and house_elections.
# Create a new column to divide the candidates into 4 groups: Democrat Not Elected,Democrat Elected, Republican Not Elected, Republican Elected.
df_candidate <- candidates %>%
  left_join(house_elections, by=c('cand_id' = 'fec_id' )) %>%
  mutate(type = ifelse(cand_party_affiliation == 'DEM', ifelse(ge_winner == 'W', 'dem_w', 'dem_nw'),ifelse(ge_winner == 'W', 'rep_w', 'rep_nw'))) %>%
  filter(!is.na(type)) %>%
  select(cand_id, type)

# Create df_contributions by joining contributions and committees. Then select columns we will need.
df_contributions <- contributions %>%
  left_join(committees, by='cmte_id') %>%
  select(cmte_id, transaction_type, transaction_amt, cand_id.x, org_type)

# Create df by joining df_candidate and df_contributions.
df <- df_candidate %>%
  left_join(df_contributions, by=c('cand_id' = 'cand_id.x' )) %>%
  arrange(cand_id) %>%
  filter(!is.na(transaction_type),!is.na(org_type), org_type !='')

# Create df1 for analysing transaction type.
df1 <- df %>%
  filter(transaction_type!='24K', transaction_type!='24C') %>%
  mutate(category=transaction_type) %>%
  group_by(type, category) %>%
  summarize(type_amt = sum(transaction_amt), type_amt_mean = mean(transaction_amt))

# Create df2 for analysing organisation type.
df2 <- df %>%
  filter(org_type!="V", org_type!="W") %>%
  mutate(category=org_type) %>%
  group_by(type, category) %>%
  summarize(type_amt = sum(transaction_amt), type_amt_mean = mean(transaction_amt)) 
```

## Creating Plots
```{r message=FALSE, echo=TRUE, warning=FALSE,}

# Function to create plots.
create_plot <- function(df = df1, is_percentage = "n", type = 'transaction', y = df1$type_amt, title = '') {
  position <- ifelse(is_percentage == 'y', 'fill', 'stack')
  plot <- df %>%
    ggplot(aes(x = type, y = y, fill= category)) +
    geom_bar(stat="identity", position=position, alpha=0.8) +
    scale_x_discrete(labels = c("Democrat\nNot Elected","Democrat\nElected","Republican\nNot Elected","Republican\nElected")) +
    theme_fivethirtyeight() +
    theme(panel.border = element_blank(), axis.title = element_blank()) +
    ggtitle(title)

  if (type=='transaction') {
    plot <- plot +
      scale_x_discrete(labels = c("Democrat\nNot Elected","Democrat\nElected","Republican\nNot Elected","Republican\nElected")) +
      scale_fill_brewer(
      palette = "Paired",
      name = "Transaction Type",
      labels = c("Independent expenditure\nopposing election", "Independent expenditure\nadvocating election", "Communication\ncost", "Communication\ncost against", "In-kind contribution made\nto registered filer"))
  } else {
    plot <- plot +
      scale_fill_manual(
      values = c("#f7c143", "#556eed", "#f2763c", "#a5a5a5"),
      name = "Organization Type",
      labels = c("Corporation", "Labor organization", "Membership organization", "Trade association"))
  }

  if (is_percentage=='y') {
    plot <- plot +
      scale_y_continuous(
        breaks = c(0, 0.25, 0.5, 0.75, 1),
        labels = c('0%', '25%', '50%','75%',  '100%'),
        expand = c(0, 0)
      )
  }
  return (plot)
}

my_plot1 <- create_plot(df = df1, is_percentage = "y", y = df1$type_amt_mean, title = "Distribution of Contributions Types in Election Campaign\n(Mean in Percentage)")

my_plot2 <- create_plot(df = df1, is_percentage = "n", y = df1$type_amt, title = "Distribution of Contributions Types in Election Campaign\n(Total in Dollar)")

my_plot3 <- create_plot(df = df2, is_percentage = "y", y = df2$type_amt_mean, type='group', title = "Distribution of Interest Groups behind Contributions\nin Election Campaign (Percentage in Dollar)")

my_plot4 <- create_plot(df = df2, is_percentage = "n", y = df2$type_amt, type='group', title = "Distribution of Interest Groups behind Contributions\nin Election Campaign (Total in Dollar)")
```
## Present Plots
```{r message=FALSE, echo=TRUE, warning=FALSE, fig.align="center", fig.width=10, fig.height=6.5}
my_plot1
my_plot2
my_plot3
my_plot4
```

##  What we can learn from these data:
The graphs presents the distibuton patterns of the contribution received by four types candaidates in 2011-2012 house election campaign, such as the findings that the contribution to republican is larger than the contribution to democrats, and the corporations are the major source for the winner candidates.


## Context / background useful in interpreting the graphic:  
For both democratcs and republicans, more money can buy more votes since campaign needs ads, mailings, and events to promote candidates.  
Because republicans support free market, lower tax, and small government, the cooperations, especially the private companies are the biggest force behind the republican who got elected.  
One article [](https://www.usnews.com/news/articles/2010/10/29/5-of-7-biggest-super-pacs-support-democrats ) suggests that 5 of 7 biggest super PACs support democrats, and some of them spending money opposing republicans' campaign, so our graph shows that the republicans no matter got elected or not elected had a large proportion of expenditure opposing them.   
Also, from the fourth graphs, the contribution from labor organizatoins to democrats is larger than it to republicans. The reason is that most labor unions favor DEM. For example, The National Education Association, the largest labor union in the United States, with a membership comprising 2.7 million public education teachers and staffers. Its super PAC has taken a targeted approach to its spending to support democrats. 

